Fail Faster

### 2025-03-12 paper/paper.tex
Add anon and review
Intro

\documentclass[sigplan,screen,acmsmall]{acmart}%%
%% \BibTeX command to typeset BibTeX logo in the docs
\AtBeginDocument{%
  \providecommand\BibTeX{{%
    Bib\TeX}}}

@@

Property-based testing (PBT) is a widely-used testing framework consisting of
two key components: a set of \textit{properties} that a system must satisfy, and
a large number of \textit{inputs} to that system. In contrast to traditional
unit testing, where inputs are hand-written, PBT users create a random-input
\textit{generator} that automatically produces inputs. Sometimes, this process
can itself be automated through the use of PBT libraries that synthesize
generators from type definitions; however, when inputs are required to maintain
invariants not communicated by the type, users must write a custom generator (or
filter out a broad class of invalid inputs).

Writing ``good" generators—those that uncover more bugs in the system under
test—is challenging. Much effort has been dedicated to this problem, often by
developing sophisticated domain-specific languages for specifying constrained
generators. However, these approaches do not always reflect the needs of real
PBT users. Under ideal conditions, a PBT suite would continue to run
until it stopped finding bugs, real-world usage patterns are very different; a
study of expert PBT users reported time budgets of between 50 milliseconds and
30 seconds for their test suites~\cite{inpractice}. Consequently, generation
time is a significant factor in PBT efficacy---that is, if a generator can
produce inputs twice as fast, it has twice as many chances within a given time
bound to find a bug. Therefore, it is imperative that PBT libraries are built
with performance in mind.

To better understand and address performance challenges in the landscape of PBT
generators, we explore popular libraries across several widely-used programming
languages. We find that the flexibility of these libraries comes at a cost to 
performance,
introducing abstraction overhead, frequent boxing and unboxing, and avoidable
allocations (which lead to costly garbage collection pauses). We focus primarily 
on OCaml's \bq, selected for its efficiency relative to other OCaml PBT libraries 
and its integration with industrial build systems, where it is expected to run 
within tight time constraints. These
inefficiencies raise a fundamental question: how can PBT libraries produce code
that is both efficient \textit{and} general? By analyzing
\bq\ and similar libraries in other languages \tr{Rust?
More?}, we provide a broader perspective on common performance challenges in PBT
and explore potential strategies for more efficient generator implementations.

Our solution is \name, a generator library that preserves
\bq’s functionality while improving performance through
\textit{multi-stage programming} (staging). Staging is a lightweight,
domain-specific compilation technique that allows us to completely eliminate the
runtime overhead of many common generator abstractions. \jwc{This relies on the fact that the code of generators is known statically at compile time} For instance, in
\name, monadic operations are zero-cost.
\name’s staged eliminates unnecessary allocations,
generates static control structures, and leverages unboxed integer libraries to
achieve performance improvements over \texttt{base\_quickcheck}, making it
well-suited to real-world PBT applications.

We present a new methodology for evaluating the relative effectiveness of PBT
generators. Prior work has compared generators based on bug-finding ability: 
generators are considered equally effective if they find bugs in a
system at the same rate over a large number of trials ~\cite{etna}. However,
because \name\ preserves the semantics of
\bq, a \name\ generator produces exactly
the same inputs as its \bq\ counterpart. By comparing
generators based on \textit{program equality}, we ensure that any speedups in
bug-finding ability stem from performance enhancements rather than variations in
input distribution. This approach establishes a foundation for benchmarking
optimizations that enhance performance without compromising expressiveness.

@@

In summary, we dramatically improve the bug-finding effectiveness of PBT generators
by optimizing them for speed. In speicific, we make the following contributions:
\begin{enumerate}
    \item An empirical analysis of the sources of inefficiency in PBT
    generators. \jwc{And an argument that PBT generator libraries should be engineered for performance}
    \item A library, \name, which offers efficient \jwc{performant?} generator \proposechange{functions}{combinators} through the
    use of multi-stage programming; this is the first known application of staging to PBT.
    \item The insight that generators should be compared by program equality 
    \item An evaluation demonstrating the improved performance of generators constructed
    using our library in a controlled comparison.

\end{enumerate}

>>>>

\documentclass[sigplan,screen,acmsmall,anonymous,review]{acmart}%%
%% \BibTeX command to typeset BibTeX logo in the docs
\AtBeginDocument{%
  \providecommand\BibTeX{{%
    Bib\TeX}}}

@@

Property-based testing (PBT) is a widely-used testing framework\hg{PBT isn't a
framework, it's a ``technique'' or ``approach'' --- Base\_QuickCheck is a framework} consisting of
two key components: a set of \textit{properties} that a system must satisfy, and
a way of producing
a large number of \textit{inputs} to that system. In contrast to traditional
unit testing, where test inputs are hand-written, PBT users create a random input
\textit{generator} that automatically produces inputs. Sometimes, this process
can itself be automated through the use of PBT libraries that synthesize
generators from type definitions; however, when inputs are required to maintain
invariants not communicated by the type, users must write a custom generator (or
filter out a broad class of invalid inputs).

\hg{I think this paragraph buries the lede. I'd start the paragraph talking
about performance and citing the 50ms number, and then jump right into what's
slow.}
Writing ``good'' generators—ones that successfully uncover bugs in the system under
test—is challenging. Much effort has been dedicated to this problem, often by
developing sophisticated domain-specific languages for specifying constrained
generators. However, these approaches do not always reflect the needs of real
PBT users. Under ideal conditions, a PBT suite would continue to run
until it stopped finding bugs, real-world usage patterns are very different; a
study of expert PBT users reported time budgets of between 50 milliseconds and
30 seconds for their test suites~\cite{inpractice}. Consequently, generation
time is a significant factor in PBT efficacy---that is, if a generator can
produce inputs twice as fast, it has twice as many chances within a given time
bound to find a bug. Therefore, it is imperative that PBT libraries are built
with performance in mind.

To better understand and address performance challenges in the landscape of PBT
generators, we explore popular libraries across several widely-used programming
languages.\hg{Again, this sentence also starts the paragraph too slowly; rather than
focus on our exploration (which isn't really the point of the paper) lead with
the result --- that abstraction overhead, boxing/unboxing, and allocations slow
down QuickCheck ports} We find that the flexibility of these libraries comes at a cost to 
performance,
introducing abstraction overhead, frequent boxing and unboxing, and avoidable
allocations (which lead to costly garbage collection pauses). We focus primarily 
on OCaml's \bq, selected for its efficiency relative to other OCaml PBT libraries 
and its integration with industrial build systems\hg{Let's not try to justify
our choice of OCaml \bq here --- we should just talk about strict functional
languages and justify this choice later}, where it is expected to run 
within tight time constraints. These
inefficiencies raise a fundamental question: how can PBT libraries produce code
that is both efficient \textit{and} general? By analyzing
\bq\ and similar libraries in other languages \tr{Rust?
More?}, we provide a broader perspective on common performance challenges in PBT
and explore potential strategies for more efficient generator implementations.

Our solution is \name, a generator library\hg{Can we say ``an approach to PBT
generator libraries'' or something more general like that?} that preserves
\bq’s functionality while improving performance through
\textit{multi-stage programming} (staging). Staging is a lightweight,
domain-specific compilation technique that allows us to completely eliminate the
runtime overhead of many common generator abstractions. \jwc{This relies on the fact that the code of generators is known statically at compile time} For instance, in
\name, monadic operations are zero-cost.
\name’s staged eliminates unnecessary allocations,
generates static control structures, and leverages unboxed integer libraries to
achieve performance improvements over \texttt{base\_quickcheck}, making it
well-suited to real-world PBT applications.

We \hg{also} present a new methodology for evaluating the relative effectiveness of PBT
generators.\hg{Might be an over-claim, we'll need to write this carefully}
Prior work has compared generators based on bug-finding ability: 
generators are considered equally effective if they find bugs in a
system at the same rate over a large number of trials ~\cite{etna}. However,
because \name\ preserves the semantics of
\bq, a \name\ generator produces exactly
the same inputs as its \bq\ counterpart. By comparing
generators based on \textit{program equality}, we ensure that any speedups in
bug-finding ability stem from performance enhancements rather than variations in
input distribution. This approach establishes a foundation for benchmarking
optimizations that enhance performance without compromising expressiveness.
\hg{I'm not totally convinced by this paragraph. Here's how I'd explain this:\\
In order to justify the switching cost associated with using \name, we designed
\name\ to be incredibly easy to benchmark: most PBT generator optimizations make
changes up to {\em distributional equivalence}, allowing the precise sampling
order to differ as long as the probability distribution is maintained, but
\name's optimizations are correct up to {\em program equivalence}. This means
that a \name\ generator produces precisely the same values in the same order as
the original generator, only faster. With this setup, benchmarking becomes
significantly easier than for prior work---our measurements have incredibly low
variance, and performance gains from \name\ translate directly to improved
bug-finding speed.}

@@

In summary, we dramatically improve the bug-finding effectiveness of PBT generators
by optimizing them for speed. Specifically, we make the following contributions:
\begin{enumerate}
    \item An empirical analysis of the sources of inefficiency in PBT
    generators. \jwc{And an argument that PBT generator libraries should be
    engineered for performance}
    \hg{Let's talk about this, I'm not sure I buy it}
    \item A library, \name, which offers efficient \jwc{performant?} generator \proposechange{functions}{combinators} through the
    use of multi-stage programming; this is the first known application of
    staging to PBT.\hg{Let's also talk about this}
    \item The insight that generators should be compared by program equality
    \item An evaluation demonstrating the improved performance of generators constructed
    using our library in a controlled comparison.
\end{enumerate}
